\emph{Light Exposure Behavior Assessment (LEBA)}: Develop of a novel instrument to capture light exposure-related behaviours

### 2021-09-04 manuscript.Rmd

```{r setup, include = FALSE}
library(papaja) # devtools::install_github("crsh/papaja")
library(lavaan)
library(semPlot)
library(semTools)
library(MOTE)
library(psych)
library(dlookr)
library(tidyverse)
library(qgraph)
library(kableExtra)
library(paran) # Parallel analysis
library(EFA.MRFA) # Hull method
library(DiagrammeR) #devtools::install_github('rich-iannone/DiagrammeR')
library(DiagrammeRsvg) #For DiagrammeR
library(rsvg) #For DiagrammeR
library(correlation)
library(ggcorrplot)
library(car) #Reversecoding
library(Hmisc)
r_refs("references.bib")
```

@@

## Merge "0"s and "1"s into "1"s 9select,subset assign)
descriptives.data[ , 6:53 ][ descriptives.data[ , 6:53 ] == 0 ] <- 1
EFA.descriptives <- descriptives.data[1:428,]
CFA.descriptives <- descriptives.data[429:690, ]

@@

### EFA data
EFA.data <- sem.data[1:428, ]

### CFA Data
CFA.data <- sem.data[429:690,]
CFARitem8 = recode(CFA.data$item8, '1=5; 2=4; 3=3; 4=2; 5=1')
CFARitem37 = recode(CFA.data$item37, '1=5; 2=4; 3=3; 4=2; 5=1')
CFA.data <- CFA.data %>% 
  mutate(Ritem8 = CFARitem8, Ritem37 = CFARitem37)

@@

```{r descriptives, eval=FALSE, include=FALSE}
#Use these  data frames for descriptive
EFA.descriptives #Only EFA
CFA.descriptives #Only CFA

@@

```{r Normality assumption EFA, include=FALSE}

@@

normality.tab = matrix(, nrow = 48, ncol = 2)
as.table(normality.tab)

rownameprefix <- "Item"
rownamesufix <- seq(1:48)
my.item.names <- paste(rownameprefix,rownamesufix, sep = "" )

rownames(normality.tab) <- my.item.names
colnames(normality.tab) <- c("statistic", "p")

@@

# Descriptives
Descriptives <- psych::describe(EFA.data)
alpha <- psych::alpha(EFA.data)
Item.total <- alpha$item.stats

@@

colnames(Des.combined) = c("Mean", "SD", "Skew", "Kurtosis", "Shapiro-Wilk Statistics", 
                           "Item-Total Correlation")

rownameprefix <- "Item"
rownamesufix <- seq(1:48)
my.names <- paste(rownameprefix,rownamesufix, sep = "" )

rownames(Des.combined) = (my.names)

#Polychoric Correlation matrix

@@

min.cor <- abs(min(upper)) #minimum cor.coefficient of the matrix
max.cor <- abs(max(upper)) # max. correlation coefficent of the matrix

# Calculating the percentange of correlations higher than .30

BigR = sum(correlations$rho >= abs(.30) & correlations$rho < abs(1.0), na.rm =T)/2 
totR = length(EFA.data)*(length(EFA.data)-1)/2
cor.per <- print (BigR/totR)*100

# Item analysis
Item_analysis <- psych::alpha(EFA.data)
Item_analysis$item.stats$r.cor
R.cor.sum <- summary(Item_analysis$item.stats$r.cor)
low.r.corec <- abs(R.cor.sum[1])  #minimum item total correlation
high.r.corec <- R.cor.sum[6]     # maximum item total correlation
```

Sampling adequacy was checked using Kaiser-Meyer-Olkin (KMO) measures of sampling adequacy[@kaiserIndexFactorialSimplicity1974] . The overall KMO vale for 48 items was `r printnum(KMO$MSA)` which was above the cutoff value (.50) indicating a mediocre sample [@hutchesonMultivariateSocialScientist1999].

@@

Table\@ref(tab:tabDes) summarizes the univariate descriptive statistics for the 48 items. some of the items were skewed with high Kurtosis values. Our data violated both univariate normality (Shapiro-Wilk statistics; [@shapiroAnalysisVarianceTest1965]) and multivariate normality assumptions (Marida's test;[@mardiaMeasuresMultivariateSkewness1970]). Multivariate skew was  = `r printnum(mardia$b1p)` (p <0.001) and multivariate kurtosis was = `r printnum(mardia$b2p)` (p <0.001). Due to these violations and ordinal nature of the response data  polychoric correlations over Pearson's correlations was chosen [@desjardinsHandbookEducationalMeasurement2018]. Bartlett's test of sphericity [@bartlettNoteMultiplyingFactors1954], $\chi^2$ (`r bartlet$df`) = `r apa(bartlet$chisq, 2, T)`, p < .001] indicated the correlations between items are adequate for the EFA. However only `r apa(cor.per, 2,T)`% of the inter-item correlation coefficients were greater than .30. The inter item correlation ranged between `r apa(min.cor,2,F)` to `r apa(max.cor,2,F)`. And the corrected item-total correlations ranged between `r apa(low.r.corec,2,F)` to `r apa(high.r.corec,2,F)`.

@@

```{r Cronbach alpha for 5 fators, message=TRUE, warning=FALSE, include=FALSE}
Five.F1 <- dplyr::select(EFA.data, item16, item36, item17)
Five.F2 <- dplyr::select(EFA.data, item11, item10, item12, item7, item8, item9) 
Five.F3 <-  dplyr::select(EFA.data, item27, item3,item40, item30, item41)
Five.F4 <-  dplyr::select(EFA.data, item33,item32,item35, item37, item38)
Five.F5 <- dplyr::select(EFA.data, item46, item45, item25, item4, item1, item26)

@@

#EFA.factor <- sapply(EFA.data, as.factor)
#correlations <- polychoric(EFA.factor, correct = 0)

@@

```{r CFA 23 items, eval=FALSE, include=FALSE}

#CFA.data <- sem.data[429:690,]

Ritem44 = recode(CFA.data$item44, '1=5; 2=4; 3=3; 4=2; 5=1')
Ritem21 = recode(CFA.data$item44, '1=5; 2=4; 3=3; 4=2; 5=1')

CFA.data <- CFA.data %>% 
  mutate(Ritem44 = Ritem44, Ritem21 = Ritem21)

LB.model.Cor <- "F1 =~ item43 + item26 + item32
                 F2 =~ item10 + item47 + item36 + Ritem44 + item35 + item13
                 F3 =~  item14 + item41+ item7 + item27 + Ritem21
                 F4 =~ item19 + item34 + item2
                 F5 =~ item5 + item24 + item30 + item1 + item16 + item28"

fit.Cor <- cfa(LB.model.Cor, data = CFA.data, estimator = "MLR", mimic = "Mplus", check.gradient = FALSE)

## Summary of Model 
summary(fit.Cor, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

## Selected Fit measures 
fit.1 <- fitmeasures (fit.Cor,c("gfi", "agfi", "nfi","rfi", 
                       "cfi.robust","tli.robust",
                       "rmsea.robust", 
                       "srmr","aic"))

semTools::reliability(fit.Cor)

@@

```{r CFA23withMI, eval=FALSE, include=FALSE}
LB.model.Cor.3 <- "F1 =~ item43 + item26 + item32
             F2 =~ item10 + item47 + item36 + item44 + item35 + item13
            F3 =~  item14 + item41+ item7 + item27 + item21
            F4 =~ item19 + item34 + item2
F5 =~ item5 + item24 + item30 + item1 + item16 + item28 
item44 ~~ item13
item14 ~~  item7
item10 ~~ item13"

## Fit the model 
fit.Cor.3 <- cfa(LB.model.Cor.3, data = CFA.data, estimator = "MLR", mimic = "Mplus")

reliability <- semTools::reliability(fit.Cor.3, return.total = T)
as.table(reliability)
## Summary of Model 
summary(fit.Cor.3, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

## Selected Fit measures 
fit.3 <- fitmeasures (fit.Cor.3,c("gfi", "agfi", "nfi","rfi", 
                       "cfi.robust","tli.robust",
                       "rmsea.robust", 
                       "srmr","aic"))

rownames <-  ("fitIndices")
colnames <- c("gfi", "agfi", "nfi","rfi", 
                       "cfi.robust","tli.robust",
                       "rmsea.robust", 
                       "srmr","aic")

@@

```{r CFI18items, eval=FALSE, include=FALSE}
LB.model.Cor.2 <- "F1 =~ item43 + item26 + item32
             F2 =~ item10 + item47 + item36 + item44 + item35 
            F3 =~  item14 + item41+ item7 
            F4 =~ item19 + item34 + item2
F5 =~ item5 + item24 + item30 + item16 
item14 ~~  item7"

#item 1 &28 discarded from F5;
#item13 discarded from F2
#item 27 & 21 discarded from F3

## Fit the model 
fit.Cor.2 <- cfa(LB.model.Cor.2, data = LB.cfa, estimator = "MLR", mimic = "Mplus")

## Summary of Model 
summary(fit.Cor.2, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

## Selected Fit measures 
fit.2 <- fitmeasures (fit.Cor.2,c("gfi", "agfi", "nfi","rfi", 
                       "cfi.robust","tli.robust",
                       "rmsea.robust", 
                       "srmr","aic"))

rownames <-  ("fitIndices")
colnames <- c("gfi", "agfi", "nfi","rfi", 
                       "cfi.robust","tli.robust",
                       "rmsea.robust", 
                       "srmr","aic")

>>>>

```{r setup, include = FALSE}
library(papaja) # devtools::install_github("crsh/papaja")
library(lavaan)
library(semPlot) #devtools::install_github('SachaEpskamp/semPlot',  dependencies = T)
library(semTools)
library(MOTE)
library(psych)
library(dlookr)
library(tidyverse)
library(qgraph)
library(kableExtra)
library(paran) # Parallel analysis
library(EFA.MRFA) # Hull method
library(DiagrammeR) #devtools::install_github('rich-iannone/DiagrammeR')
library(DiagrammeRsvg) #For DiagrammeR
library(rsvg) #For DiagrammeR
library(correlation)
library(ggcorrplot)
library(car) #Reversecoding
library(Hmisc)
r_refs("references.bib")
```

@@

## Merge "0"s and "1"s into "1"s 9select,subset assign)
descriptives.data[ , 6:53 ][ descriptives.data[ , 6:53 ] == 0 ] <- 1
EFA.descriptives <- descriptives.data[1:428,]
CFA.descriptives <- descriptives.data[429:690, ]
IRT.descriptives <- descriptives.data

@@

@@

```{r descriptives, eval=FALSE, include=FALSE}
#Use these  data frames for descriptive
EFA.descriptives #Only EFA, n =428
CFA.descriptives #Only CFA, n =262  
IRT.descriptives# Total sample, n = 690

@@

```{r EFAdata, include=FALSE}
### EFA data
EFA.data <- sem.data[1:428, ]
```

```{r EFAassumptions, include=FALSE}

@@

normality.tab = matrix(, nrow = 48, ncol = 2)
as.table(normality.tab)
rownameprefix <- "Item"
rownamesufix <- seq(1:48)
my.item.names <- paste(rownameprefix,rownamesufix, sep = "" )
rownames(normality.tab) <- my.item.names
colnames(normality.tab) <- c("statistic", "p")

@@

# Descriptive Stats
Descriptives <- psych::describe(EFA.data)
alpha <- psych::alpha(EFA.data)
Item.total <- alpha$item.stats

@@

colnames(Des.combined) = c("Mean", "SD", "Skew", "Kurtosis", "Shapiro-Wilk Statistics","Item-Total Correlation")

rownameprefix <- "Item"
rownamesufix <- seq(1:48)
my.names <- paste(rownameprefix,rownamesufix, sep = "" )
rownames(Des.combined) = (my.names)
```

```{r CorrelationMatrix, include=FALSE}
#Polychoric Correlation matrix

@@

min.cor <- abs(min(upper)) #minimum cor.coefficient of the matrix
max.cor <- abs(max(upper)) # max. correlation coefficient of the matrix

# Calculating the percentage of correlations higher than .30

BigR = sum(correlations$rho >= abs(.30) & correlations$rho < abs(1.0), na.rm =T)/2 
totR = length(EFA.data)*(length(EFA.data)-1)/2
cor.per <- print (BigR/totR)*100
```

```{r ItemAnalysis, include=FALSE}
# Item analysis
Item_analysis <- psych::alpha(EFA.data)
Item_analysis$item.stats$r.cor
R.cor.sum <- summary(Item_analysis$item.stats$r.cor)
low.r.corec <- abs(R.cor.sum[1])  #minimum item total correlation
high.r.corec <- R.cor.sum[6]     # maximum item total correlation
```

Sampling adequacy was checked using Kaiser-Meyer-Olkin (KMO) measures of sampling adequacy[@kaiserIndexFactorialSimplicity1974] . The overall KMO vale for 48 items was `r printnum(KMO$MSA)` which was above the cutoff value (.50) indicating a mediocre sample [@hutchesonMultivariateSocialScientist1999]. Table\@ref(tab:tabDes) summarizes the univariate descriptive statistics for the 48 items. some of the items were skewed with high Kurtosis values. Our data violated both univariate normality (Shapiro-Wilk statistics; [@shapiroAnalysisVarianceTest1965]) and multivariate normality assumptions (Marida's test;[@mardiaMeasuresMultivariateSkewness1970]). Multivariate skew was  = `r printnum(mardia$b1p)` (p <0.001) and multivariate kurtosis was = `r printnum(mardia$b2p)` (p <0.001). Due to these violations and ordinal nature of the response data  polychoric correlations over Pearson's correlations was chosen [@desjardinsHandbookEducationalMeasurement2018]. Bartlett's test of sphericity [@bartlettNoteMultiplyingFactors1954], $\chi^2$ (`r bartlet$df`) = `r apa(bartlet$chisq, 2, T)`, p < .001] indicated the correlations between items are adequate for the EFA. However only `r apa(cor.per, 2,T)`% of the inter-item correlation coefficients were greater than .30. The inter item correlation ranged between `r apa(min.cor,2,F)` to `r apa(max.cor,2,F)`. And the corrected item-total correlations ranged between `r apa(low.r.corec,2,F)` to `r apa(high.r.corec,2,F)`.

@@

@@

```{r Cronbach alpha for 5 fators, message=TRUE, warning=FALSE, include=FALSE}

Five.F1 <- dplyr::select(EFA.data, item16, item36, item17)
Five.F2 <- dplyr::select(EFA.data, item11, item10, item12, item7, item8, item9) 
Five.F3 <-  dplyr::select(EFA.data, item27, item3,item40, item30, item41)
Five.F4 <-  dplyr::select(EFA.data, item33,item32,item35, item37, item38)
Five.F5 <- dplyr::select(EFA.data, item46, item45, item25, item4, item1, item26)

@@

@@

```{r CFA data frame, include=FALSE}

### CFA Data
CFA.data <- sem.data[429:690,]
CFARitem8 = recode(CFA.data$item8, '1=5; 2=4; 3=3; 4=2; 5=1')
CFARitem37 = recode(CFA.data$item37, '1=5; 2=4; 3=3; 4=2; 5=1')
CFARitem26 = recode(CFA.data$item26, '1=5; 2=4; 3=3; 4=2; 5=1')
CFA.data <- CFA.data %>% 
  mutate(Ritem8 = CFARitem8, Ritem37 = CFARitem37, Ritem26 = CFARitem26)
```

```{r CFA 25items, include=FALSE }

LB.model.Cor.1 <- "F1 =~ item16 + item36 + item17
                 F2 =~ item11 + item10 + item12+ item7+ Ritem8+ item9
                 F3 =~  item27+ item3+ item40 + item30 + item41
                 F4 =~ a1*item32 +item35 + Ritem37 + item38+ a2*item33 
                 F5 =~ item46+ item45+ item25+ item4+ item1+ item26

a1==a2"

fit.Cor.1 <- cfa(LB.model.Cor.1, data = CFA.data, ordered = names(CFA.data),estimator = "WLSMV")

## Summary of Model 
summary(fit.Cor.1, fit.measures =TRUE,standardized = TRUE)

## Selected Fit measures 
fit.1 <- fitmeasures (fit.Cor.1,c("gfi", "agfi", "nfi","rfi", 
                       "cfi.scaled","tli.scaled",
                       "rmsea.robust", 
                       "srmr","aic"))

semTools::reliability(fit.Cor.1)

@@

```{r mod1, include=FALSE}
modfit.Cor.one <- modindices(fit.Cor.1, sort. = TRUE) 
modfit.Cor.one[modfit.Cor.one$mi>3.84,]
```

```{r CFA23withMI, include=FALSE}
LB.model.Cor.2 <- "F1 =~ item16 + item36 + item17
                 F2 =~ item11 + item10 + item12+ item7+ Ritem8+ item9
                 F3 =~  item27+ item3+ item40 + item30 + item41
                 F4 =~ a1*item32 +item35 + item38+ a2*item33 
                 F5 =~ item46+ item45+ item25+ item4+ item1
                       a1==a2 
                item30 ~~  item41"

#Ritem26 & 37 are removed

## Fit the model 
fit.Cor.2 <- cfa(LB.model.Cor.2, data = CFA.data, ordered = names(CFA.data),estimator = "WLSMV")

reliability <- semTools::reliability(fit.Cor.2, return.total = T)
as.table(reliability)
## Summary of Model 
summary(fit.Cor.2, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

## Selected Fit measures 
fit.2 <- fitmeasures (fit.Cor.2,c("gfi", "agfi", "nfi","rfi", 
                       "cfi.scaled","tli.scaled",
                       "rmsea.scaled", 
                       "srmr","aic"))

@@

```{r mod2, include=FALSE}
modfit.Cor.one <- modindices(fit.Cor.2, sort. = TRUE) 
modfit.Cor.one[modfit.Cor.one$mi>3.84,]
```

```{r figcfa, fig.cap = "(A) Five Factor Model of LEBA", , out.width="100%",  warning=FALSE}
library(semPlot)

semPaths (fit.Cor.2 , 
          what= "std",
          #whatLabels = "std",
          intercepts = F,
          #style ="lisrel",
          #residScale = 6,
          theme = "colorblind",
          nCharNodes = 0,
          reorder =T,
          rotation =2,
          layout ="tree",
          cardinal = T,
          curvePivot =T,
          sizeMan =8,
          sizeMan2 = 1.5,
          sizeLat = 8,
          #edge.color = "black",
          thresholds = FALSE,
          equalizeManifests =T,
          fade = FALSE,
          edge.label.cex = .65,
          exoCov = FALSE)
